Cost distance inputs

### 2016-02-14 Rmd/all-inputs.Rmd
Many updates. Up through new analysis.

crops <- c("maize", "soy")
ynms <- dir(p_yield, pattern = "fao")  # FAO adjusted now
ylds <- lapply(ynms, function(x) brick(full_path(p_yield, x)))
names(ylds) <- crops
ylds <- stack(list(ylds[[1]][[5]], ylds[[2]][[4]]))  # take the last two of each
names(ylds) <- crops
ylds <- round(ylds, 2)

@@

## Stack and mask all grids
```{r, eval = FALSE}
# deal with yields
# potmaize <- ylds$maize_R  # just taking rainfed commercial potential to start
# w <- matrix(nrow = 5, ncol = 5, rep(1, 5 * 5))  # have to filter b/c artifacts
# potmaizef <- focal(potmaize[[1]], w = w, fun = mean, na.rm = TRUE)
# potsoy <- ylds$soybean_R  # just taking rainfed commercial potential to start
# potsoy[potsoy > 5] <- 5 #round(cellStats(potsoy, mean), 1)  # really high fix
# potsoyf <- focal(potsoy[[1]], w = w, fun = mean, na.rm = TRUE)
# # potsoy <- mask(potsoyf, potsoy)  # crop yield
# 
yield_fix <- function(x, w, na.rm = TRUE) {
  r <- focal(x, w = w, fun = mean, na.rm = na.rm)
  rmsk <- (!is.na(r) & is.na(x)) * r
  x[is.na(x)] <- 0
  rfix <- x + rmsk  # add back in edge bits
}
w <- matrix(nrow = 9, ncol = 9, rep(1, 81))
mzf <- yield_fix(ylds[[1]], w)
soyf <- yield_fix(ylds[[2]], w)
ylds_s <- stack(sapply(list(mzf, soyf), function(x) {
  x[is.na(x)] <- 0
  x
}))
ylds_s <- raster::mask(ylds_s, mgrid)
names(ylds_s) <- crops
# plot(ylds_s[[1]])
# potmaizef2 <- yield_fix(potmaizef[[1]], w)
# plot(potmaizef2)
# # first get rid of all NA area in all grids
# # start with yield grids
# # yldmask <- calc(stack(ylds), sum)
# 
# # Function to disaggregate and reproject yield rasters
# yld_disagg <- function(x, msk, fact = 10) {
#   r <- disaggregate(x, fact = 10)
#   r2 <- resample(x, r)
#   r3 <- projectRaster(r2, msk)
#   r4 <- mask(r3, msk)
#   r4
# }
# 
# ylds_1k <- lapply(list(potmaizef2, potsoyf2), function(x) {
#   yld_disagg(x, msk = mgrid)
# })
# ylds_s <- stack(ylds_1k) 
# names(ylds_s) <- c("maize", "soy")
# ylds_s <- round(ylds_s, 1)  
# plot(ylds_s)

@@

# pas and intactness
intpa <- stack(intact, pas)

mgridc <- crop(mgrid, avail)
inlist <- list("convertible" = avail, "bd" = bd, "carbon" = carbonf, 
               "cost" = cost, "cons" = cons, "cropland" = cropland, 
               "intpa" = intpa, "p_yield" = ylds_s)
inlist <- lapply(inlist, function(x) crop(x, mgridc))
inlist2 <- lapply(inlist, function(x) calc(x, mean))
inlist3 <- calc(stack(inlist2), mean)
# inlist4 <- calc(stack(inlist3, crop(ylds_s, mgridc)), mean)

@@

# write out mask raster
base <- as.data.table(msk, xy = TRUE)
setnames(base, old = names(msk), new = "val")
base[, ind := 1:nrow(base)]
base <- base[!is.na(val)][, val := NULL]
fnm <- full_path(p_production, paste0(cnames, "-mask.csv"))
write.table(base, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
# plot(dt_to_raster(base, mgridf@crs))

@@

# Copy over files to the correct directory
```{r, eval}
p_mod <- full_path(p_proj, "agroEcoTradeoff/external/data/ZA/")
p_master <- "~/Dropbox/projects/agroEcoTradeoff/external/data/ZA/"
fnms <- dir(p_production)
file.copy(dir(p_production, full.names = TRUE), full_path(p_mod, fnms),
          overwrite = TRUE)
file.copy(dir(p_production, full.names = TRUE), full_path(p_master, fnms),
          overwrite = TRUE)

>>>>

crops <- c("maize", "soy")
ynms <- dir(p_yield, pattern = "fao")  # FAO adjusted now
ylds <- lapply(ynms, function(x) brick(full_path(p_yield, x)))
names(ylds) <- crops
ylds <- stack(list(ylds[[1]][[1]], ylds[[2]][[1]]))  # take first two of each
names(ylds) <- crops
ylds <- round(ylds, 2)

@@

## Stack and mask all grids
```{r, eval = FALSE}
# deal with yields
# potmaize <- ylds$maize_R  # just taking rainfed commercial potential to start
# w <- matrix(nrow = 5, ncol = 5, rep(1, 5 * 5))  # have to filter b/c artifacts
# potmaizef <- focal(potmaize[[1]], w = w, fun = mean, na.rm = TRUE)
# potsoy <- ylds$soybean_R  # just taking rainfed commercial potential to start
# potsoy[potsoy > 5] <- 5 #round(cellStats(potsoy, mean), 1)  # really high fix
# potsoyf <- focal(potsoy[[1]], w = w, fun = mean, na.rm = TRUE)
# # potsoy <- mask(potsoyf, potsoy)  # crop yield
# 
yield_fix <- function(x, w, na.rm = TRUE) {
  r <- focal(x, w = w, fun = mean, na.rm = na.rm)
  rmsk <- (!is.na(r) & is.na(x)) * r
  x[is.na(x)] <- 0
  rfix <- x + rmsk  # add back in edge bits
  # rfix[rfix == 0] <- NA
  # rfix
}
w <- matrix(nrow = 9, ncol = 9, rep(1, 81))
mzf <- yield_fix(ylds[[1]], w)
soyf <- yield_fix(ylds[[2]], w)
yld_s <- stack(list(mzf, soyf))
# ylds_s <- stack(sapply(list(mzf, soyf), function(x) {
#   x[is.na(x)] <- 0
#   x
# }))
yld_s <- raster::mask(yld_s, mgrid)
names(yld_s) <- crops
# plot(yld_s[[1]])
# potmaizef2 <- yield_fix(potmaizef[[1]], w)
# plot(potmaizef2)
# # first get rid of all NA area in all grids
# # start with yield grids
# # yldmask <- calc(stack(ylds), sum)
# 
# # Function to disaggregate and reproject yield rasters
# yld_disagg <- function(x, msk, fact = 10) {
#   r <- disaggregate(x, fact = 10)
#   r2 <- resample(x, r)
#   r3 <- projectRaster(r2, msk)
#   r4 <- mask(r3, msk)
#   r4
# }
# 
# ylds_1k <- lapply(list(potmaizef2, potsoyf2), function(x) {
#   yld_disagg(x, msk = mgrid)
# })
# ylds_s <- stack(ylds_1k) 
# names(ylds_s) <- c("maize", "soy")
# ylds_s <- round(ylds_s, 1)  
# plot(ylds_s)

@@

# pas and intactness
intpa <- stack(intact, pas)
mgridc <- crop(mgrid, avail)
inlist <- list("convertible" = avail, "bd" = bd, "carbon" = carbonf, 
               "cost" = cost, "cons" = cons, "cropland" = cropland, 
               "intpa" = intpa, "p_yield" = yld_s)
inlist <- lapply(inlist, function(x) crop(x, mgridc))
inlist2 <- lapply(inlist, function(x) calc(x, mean))
inlist3 <- calc(stack(inlist2), mean)
# inlist4 <- calc(stack(inlist3, crop(ylds_s, mgridc)), mean)

@@

# write out mask raster
base <- as.data.table(msk, xy = TRUE)
setnames(base, old = names(msk), new = "val")
base[, ind := 1:nrow(base)]
base <- base[!is.na(val)][, val := NULL]
fnm <- full_path(p_production, paste0(cnames, "-mask.csv"))
write.table(base, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
# plot(dt_to_raster(base, mgridf@crs))
# plot(dt_to_raster(cbind(base[, .(x, y)], DT), mgridf@crs))

@@

# Copy over files to the correct directory
```{r, eval}
p_mod <- full_path(p_proj, "agroEcoTradeoff/external/data/ZA/")
p_master <- "~/Dropbox/projects/agroEcoTradeoff/external/data/ZA/"
p_downloads <- "~/Dropbox/projects/agroEcoTradeoff/external/data/downloads/"
fnms <- dir(p_production)
file.copy(dir(p_production, full.names = TRUE), full_path(p_mod, fnms),
          overwrite = TRUE)
file.copy(dir(p_production, full.names = TRUE), full_path(p_master, fnms),
          overwrite = TRUE)
setwd(p_production)
zip("demodat.zip", dir())
file.copy("demodat.zip", fp(p_downloads, "demodat.zip"), overwrite = TRUE)
file.remove("demodat.zip")
setwd(p_proj)

### 2016-02-14 Rmd/biodiversity-layer.Rmd
Many updates. Up through new analysis.

## Prepare impacts layer
```{r, eval = FALSE}
vnms <- dir(full_path(p_bd, "act_veg"), pattern = ".tif$")
vnms <- vnms[vnms != "veg88.tif"]  # drop water bodies
fnms <- full_path(full_path(p_bd, "act_veg"), vnms)
s <- stack(lapply(fnms, function(x) raster(x)))
sm <- mask(s, mgrid)

>>>>

## Prepare impacts layer
```{r, eval = FALSE}
vnms <- dir(full_path(p_bd, "act_veg"), pattern = ".tif$")
vnms <- vnms[vnms != "veg88.tif"]  # drop water bodies
fnms <- full_path(full_path(p_bd, "act_veg"), vnms)
s <- stack(lapply(fnms, function(x) raster(x)))
sm <- raster::mask(s, mgrid)

### 2016-02-14 Rmd/cost-distance.Rmd
Many updates. Up through new analysis.

```{r, eval = FALSE}
library(lmisc)
library(raster)
library(rgdal)
library(rgeos)
library(gdalUtils)
p_root <- lmisc::set_base_path("zambiaToff")
p_roads <- full_path(p_root, "external/input_devel/roads/")
p_roads2 <- paste0("~/Dropbox/data/zari/infrastructure/zambia/rda_dump/", 
                   "Final Shape files RDA/")
p_trade <- "~/Dropbox/projects/agroEcoTradeoff/external/base_data/"
p_cost <- full_path(p_root, "external/input_devel/roads/cost")

@@

# towns
setwd(p_roads)
ztowns <- readOGR(dsn = "district_towns.shp", layer = "district_towns")
ntowns <- readOGR("zambia-neighbor-cities.shp", 
                  layer = "zambia-neighbor-cities")
ntowns <- ntowns[ntowns$ES00POP > 10000, ]  # filter down to 10,000 or greater
ztowns$cid <- 1:nrow(ztowns)
ztowns@data <- ztowns@data[, "cid", drop = FALSE]  # reduce for binding
ntowns$cid <- (max(ztowns$cid) + 1):(nrow(ntowns) + (max(ztowns$cid)))
ntowns@data <- ntowns@data[, "cid", drop = FALSE]  # reduce for binding
ntowns@proj4string <- ztowns@proj4string
alltowns <- spTransform(rbind(ztowns, ntowns), CRSobj = projection(zfr))

# write this out--needed for determining nodes in cost distance analysis
writeOGR(alltowns, dsn = "alltowns.shp", layer = "alltowns", 
         driver = "ESRI Shapefile", overwrite = TRUE)

@@

roadnms <- sapply(rtypes, function(x) {
  dir(p_roads, pattern = paste0("^roads_", x))
})
rdrcl <- c(1, 2, 4, 8)
roads <- lapply(1:length(roads), function(x) {
  nm <- unname(roadnms)[x]
  rdspoly <- readOGR(nm, layer = gsub("\\.sqlite", "", nm))
  rdspoly@proj4string <- zframe@proj4string
  rdspoly$code <- rdrcl[x]  # assign new variable with code for road coding
  rdspoly@data <- rdspoly@data[, c("full_id", "osm_id", "code")]
  rdspoly
})

@@

# plot(zframe)
# plot(lakes, add = TRUE)
# plot(ztowns, pch = 20, cex = 0.5)
# points(ntowns, pch = 20, cex = 0.5)

@@

>>>>

```{r, eval = FALSE}
library(lmisc)
library(raster)
library(rgdal)
library(rgeos)
library(gdalUtils)
setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- lmisc::set_base_path("zambiaToff")
p_roads <- full_path(p_root, "external/input_devel/roads/")
p_roads2 <- paste0("~/Dropbox/data/zari/infrastructure/zambia/rda_dump/", 
                   "Final Shape files RDA/")
p_trade <- "~/Dropbox/projects/agroEcoTradeoff/external/base_data/"
p_cost <- full_path(p_root, "external/input_devel/roads/cost")

@@

# towns
setwd(p_roads)
# ztowns <- readOGR(dsn = "district_towns.shp", layer = "district_towns")
# ntowns <- readOGR("zambia-neighbor-cities.shp", 
#                   layer = "zambia-neighbor-cities")
# ntowns <- ntowns[ntowns$ES00POP > 10000, ]  # filter down to 10,000 or greater
# ztowns$cid <- 1:nrow(ztowns)
# ztowns@data <- ztowns@data[, "cid", drop = FALSE]  # reduce for binding
# ntowns$cid <- (max(ztowns$cid) + 1):(nrow(ntowns) + (max(ztowns$cid)))
# ntowns@data <- ntowns@data[, "cid", drop = FALSE]  # reduce for binding
# ntowns@proj4string <- ztowns@proj4string
# alltowns <- spTransform(rbind(ztowns, ntowns), CRSobj = projection(zfr))

# write this out--needed for determining nodes in cost distance analysis
# writeOGR(alltowns, dsn = "alltowns.shp", layer = "alltowns", 
#          driver = "ESRI Shapefile", overwrite = TRUE)

alltowns <- readOGR("alltowns.shp", layer = "alltowns")

@@

roadnms <- sapply(rtypes, function(x) {
  dir(p_roads, pattern = paste0("^roads_", x))
})
rdrcl <- c(1, 2, 4, 8)
roads <- lapply(1:length(roadnms), function(x) {
  nm <- unname(roadnms)[x]
  rdspoly <- readOGR(nm, layer = gsub("\\.sqlite", "", nm))
  rdspoly@proj4string <- zframe@proj4string
  rdspoly$code <- rdrcl[x]  # assign new variable with code for road coding
  rdspoly@data <- rdspoly@data[, c("full_id", "osm_id", "code")]
  rdspoly
})

@@

tmd2 <- spChFIDs(tmd, as.character(1:nrow(tmd)))
pfr2 <- spChFIDs(pfr, as.character((nrow(tmd) + 1):(nrow(tmd) + nrow(pfr))))
tmd2@data <- tmd2@data[, "roadid", drop = FALSE]
pfr2@data <- pfr2@data[, "roadid", drop = FALSE]

zroads <- rbind(tmd2, pfr2)
zroads <- spTransform(zroads, mgrid@crs)
# save(zroads, file = fp(p_roads, "tmdpfroads.rda"))
# plot(zframe)
# plot(lakes, add = TRUE)
# plot(ztowns, pch = 20, cex = 0.5)
# points(ntowns, pch = 20, cex = 0.5)

@@

## Later fix

Attempt to fix roads so they display properly in model output maps. Essentially combined OSM roads with Zambia roads, rasterized them to 30 m, added them, reduced all to values of 1, and then within ArcGIS (a bit of it earlier done in QGIS) I converted to polylines. Results used in main-analysis.Rmd. 
```{r, eval = FALSE}
# roads
setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- getwd()
p_demand <- fp(p_root, "external/input_devel/demand/")
p_mod <- fp(p_root, "agroEcoTradeoff")
p_moddat <- fp(p_mod, "external/data/ZA")
p_figures <- full_path(p_root, "paper/figures/")
p_roads <- fp(p_root, "external/input_devel/roads")
p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                   "rda_dump/Final Shape files RDA/")

mgrid <- raster(fp(p_moddat, "ZA-mask.tif"))
osmroads <- readOGR(fp(p_roads, "osmroads/zamroads.shp"), layer = "zamroads")

osmroads$val <- 1
tmd <- readOGR(fp(p_roads2, "TMDLinks_Merged.shp"), layer = "TMDLinks_Merged")
pfr <- readOGR(fp(p_roads2, "PFRLinks.shp"), layer = "PFRLinks")
pfrf <- pfr[-which(duplicated(pfr@data$RoadId)), ]
tmdf <- tmd[-which(duplicated(tmd@data$RoadId)), ]
tmdf <- spChFIDs(tmdf, as.character(1:nrow(tmdf)))
pfrf <- spChFIDs(pfrf, as.character((nrow(tmdf) + 1):(nrow(tmdf) + nrow(pfrf))))
tmdf@data <- tmdf@data[, "RoadId", drop = FALSE]
pfrf@data <- pfrf@data[, "RoadId", drop = FALSE]
roads <- rbind(tmdf, pfrf)
roads <- spTransform(roads, pas@proj4string)
osmroads <- spTransform(osmroads, pas@proj4string)
roads$val <- 1
plot(roads)
plot(tmd)
plot(pfr, add = TRUE, col = "red")
plot(osmroads)
plot(tmdf)
writeOGR(roads, dsn = fp(p_roads, "osmroads/tmdpfrf.shp"), layer = "tmpdpfr", 
         driver = "ESRI Shapefile", overwrite_layer = TRUE)
writeOGR(osmroads, dsn = fp(p_roads, "osmroads/zamroadsalb.shp"), 
         layer = "zamroadsalb", driver = "ESRI Shapefile")
writeOGR(spTransform(roads, osmroads@proj4string),
         dsn = fp(p_roads, "tmdpfrf_gcs.shp"), layer = "tmpdpfr", 
         driver = "ESRI Shapefile")
inm <- fp(p_roads, "osmroads/zamroadsalb.shp")
onm <- fp(p_roads, "osmroads/zamroadsalb.tif")
gdal_rasterize(src_datasource = inm, l = "zamroadsalb", ot = "Byte", 
               a = "val", tr = c(30, 30), at = TRUE, co = "COMPRESS=LZW",
               dst_filename = onm, te = bbox(mgrid)[1:4], of = "GTiff", 
               verbose = TRUE)
inm <- fp(p_roads, "osmroads/zamroadsalb.shp")
onm <- fp(p_roads, "osmroads/zamroadsalb.tif")
gdal_rasterize(src_datasource = inm, l = "zamroadsalb", ot = "Byte", 
               a = "val", tr = c(30, 30), at = TRUE, co = "COMPRESS=LZW",
               dst_filename = onm, te = bbox(mgrid)[1:4], of = "GTiff", 
               verbose = TRUE)
inm <- fp(p_roads, "osmroads/tmdpfrf.shp")
onm <- fp(p_roads, "osmroads/tmdpfrf.tif")
gdal_rasterize(src_datasource = inm, l = "tmdpfrf", ot = "Byte", 
               a = "val", tr = c(30, 30), at = TRUE, co = "COMPRESS=LZW",
               dst_filename = onm, te = bbox(mgrid)[1:4], of = "GTiff", 
               verbose = TRUE)
r <- raster(fp(p_roads, "osmroads/tmdpfr_plus_osm.tif"))
r2 <- rast_math(r > 0, tmpdir = fp(p_roads, "osmroads"))
